Are those eggs in your pocket, or are you just happy to see me?

South Africa, Xiaomei Havard, ANC
===
\begin{quote}
I hoped to write a profile on this intriguing new parliamentarian — but when her responses came back, I realised it was a non-starter.

This was because the nature of her answers, due mainly to the language barrier, would have exposed her to — I felt — unjustified public derision. To give one relatively tame example, I asked her about the inclusion of her name on the list of MP members called to appear before the ANC’s Integrity Commission in 2020. This was Havard’s written response:
    \begin{quote}
        One day I found a list of 23 people was published for the Integrity Commission, also with my name there. I was so excited at that time. I did not understand what is the Integrity Commission’s role? Just saw my name was mentioned with many famous leaders. I also felt lucky to get a free Advertisement with so many respect high profile leaders, I had waited for over 2-3 weeks for appearance in front of the Integrity Commission. I also phoned the Integrity Commission if I need to appear in the Integrity Commission? The answer is no, I do not need to appear at the Integrity Commission, because I had passed Vetting.
    \end{quote}
\end{quote}
\begin{quote}
The rest of the responses were even more surreal, and included an anecdote about Havard’s late husband never leaving the house without filling his pockets with boiled eggs to distribute to the poor.
\end{quote}
\nocite{davis2021}