Letter to <em>Jom</em>

Singapore, Israel, antisemitism, Jom, Ang Swee Chai, Palestine, Hamas, 7 October
===
An old letter.
\begin{quote}
Jom’s report on Dr Ang’s presentation at the Singapore Island Country Club originally omitted to mention that in 2014 Dr Ang shared an antisemitic video about “CNN, Goldman Sachs, and the Zio Matrix” by the neo-Nazi leader of the Klu Klux Klan, David Duke. This raises the question of what exactly Dr Ang thinks “actually happened on 7 October 2023”, and whether the content of the link amounted to conspiracy theorising.

I do not mention this to detract from the quite extraordinary courage Dr Ang has displayed in her humanitarian work with Medical Aid for Palestinians, and I am glad that there is some space in Singapore to openly discuss the Israel–Hamas war and the wider conflict.

Rather, the incident serves as a reminder that the historically pro-Israeli leanings of Singaporean public and elite opinion and the general paucity of antisemitism in Singaporean history are at best imperfect barriers to antisemitism. Most people of good will, of whatever opinion, seek to distinguish and separate antisemitism from reasoned accusations against the Israeli government. (By “reasoned” I do not mean “moderate”—Jom’s piece on accusations of genocide is a good example.) To effectively make such a distinction, it is not enough to syllogistically repeat that antisemitism is racism and therefore bad; critical evaluation of whether particular material contains antisemitic ideas is necessary. Of that, I am sure that Singaporeans will prove most capable.
\end{quote}

\nocite{loo2024a}