Rory Stewart in the <em>TLS</em> on Afghanistan and plans.

Afghanistan, Times Literary Supplement, plans
===
\begin{quote}
    So when Rumsfeld says he has no idea what the US is doing, or what the mission is, or “who the bad guys are”, he does not mean, as Whitlock implies, that nobody has been producing plans or objectives or analysis. He means that he doesn’t like what he sees. The same is true when others later say that Rumsfeld himself did not have a plan. Or when various generals arrive claiming there is no plan, write a detailed plan, and then hand over to a successor who claims there is no plan, before writing a detailed plan and handing over to a successor feigning bafflement.

    But what if the interviewees and Craig Whitlock are telling you more about themselves than about Afghanistan? What if there was indeed deep corruption in Afghanistan, but that there were also highly intelligent and dedicated public servants, including President Ghani himself? What if there had never been a conscious conspiracy to conceal the problems of Afghanistan? What if the US could have kept 2,500 troops relatively safely and indefinitely in Afghanistan (as it keeps 25,000 troops, seventy years after the cessation of formal conflict, in South Korea)? What if the sunk costs of the intervention were immense, but the marginal costs of remaining were minimal? What if no magical plan could have existed for Afghanistan but, nevertheless, keeping a very few troops would have saved the country from a Taliban government, protected basic political and civil rights, supported women – at least in urban areas and much of the rural Centre and North – prevented the economy and the health and education systems from collapsing, and kept millions from the edge of starvation?

    What if the US withdrawal was determined not by the “Bloody Battle in Affghanistan” but by the “Grand Contested Election for the Presidency of the United States”? What if Afghans were not cowards or incompetents but instead as smart and good as you and me?
\end{quote}

\nocite{stewart}