Plodwatch I.

Britain, plodwatch
===
24 July 2022: Drew Pavlou (a rather unpleasant man) is \href{https://www.theguardian.com/australia-news/2022/jul/23/australian-activist-drew-pavlou-arrested-in-london-but-denies-sending-chinese-embassy-bomb-threat}{arrested} after the Chinese embassy sends itself a bomb threat and puts Pavlou’s name on it.

18 May 2025: Peter Tatchell is \href{https://www.petertatchellfoundation.org/peter-tatchell-arrested-at-palestine-protest-on-fake-charges/}{arrested} for his sign reading ‘STOP Israel Genocide! Stop Hamas executions! Odai Al-Rubai, aged 22, executed by Hamas! RIP!’ and/or shouting ‘Hamas are terrorists’ (which he says he didn’t shout, but of course there’s nothing wrong with shouting what is, incidentally, the \href{https://www.legislation.gov.uk/uksi/2021/1318/article/2/made}{position} of His Majesty’s Government).

17 July 2025: A protester holding a Palestinian flag whose signs read ‘Free Gaza’ and ‘Israel is committing genocide’ is threatened with arrest under section 12(1A) of the Terrorism Act and made to give her details to avoid arrest. 
It is quite transparent, in reading the transcript, that the officers concerned have decided that they must have the details of the person concerned and have them leave, and they will have to concoct some legal justification for that if necessary. Each of these justifications is implausible.
\begin{enumerate}
    \item They claim that an offence is being committed under the Terrorism Act. The officers variously suggest that Murton supported a proscribed group, did something that ‘come[s] under proscribed groups’ (whatever that means), and ‘express[ed] an opinion or belief that is supportive of a proscribed organisation’. None of this is true.
    \item They also suggest that Murton was distracting traffic, which is supposedly another offence. (What precisely that offence is is unclear. Presumably any advertisement is capable of distracting traffic. So too would the gathering of officers.)
    \item And apparently the officers’ cognitive capacities sufficed to suggest a ‘public order offence’ might have been committed as well.
\end{enumerate}
None of these justifications make the slightest bit of sense. Instead of answering Murton’s perfectly reasonable questions in response to them, they simply resorted to finding other (poor) excuses, or threatening arrest.

Subsequently, Kent Police told \href{https://www.thenational.scot/news/25321245.disturbed-mp-hits-armed-police-threaten-gaza-protester/}{\emph{The National}} that---
\begin{quote}
    Following the behaviour of an individual on a traffic roundabout in Canterbury on Monday, July 14, 2025, officers attended to investigate. Having ascertained no offences had been committed, no further action was taken
\end{quote}

This is surely inaccurate. The officers state clearly that at least two offences were committed, in ‘distracting’ traffic, and ‘express[ing] an opinion or belief that is supportive of a proscribed organisation’.

Here is a transcript (mostly copied from YouTube’s)—
\begin{quote}
    What's your intention here today?
    
    My intention uh is to wave this flag and keep Palestine in the public consciousness right now.
    
    So, do you support any proscribed group?
    
    I do not support any proscribed group.
    
    Okay.
    
    I support a free Palestine and the end of genocide.
    
    Okay. Can I get your details?
    
    Am I required to give them to you?
    
    Well, you may be committing [an] offence at the moment. So, I just need to make sure that you're legit.
    
    What offence?
    
    Well, as you're aware, obviously [inaudible] an offence to support a proscribed group [inaudible] Palestine Action.
    
    Yeah, but I don't have anything on which says that.
    
    Appreciate that. But the way you are behaving at the moment would lead me to believe that you maybe give me suspicion to believe grounds you could be. [sic]
    
    What suspicion? That I've got a sign that says free Gaza? Holding a Palestinian flag? and I have a sign that says Israel is committing genocide? Something which is by the way—
    
    You’re also distracting the traffic which may cause an accident. That's also another offence you're committing.
    
    Right. Okay.
    
    So I've got a reason to take your name.
    
    Do I have to give it?
    
    Potentially committing a public order offence as well.
    
    What do you mean potentially? I don't understand. Either I am or I'm not.
    
    Right, we can go about this two ways, right? My colleagues explained that the support of this, mentioning freedom of Gaza, Israel genocide, all of that all come under proscribed groups which are terror groups that have been dictated by the government. All right, which we have suspicions to suggest that you're supporting based on your actions here in a city with the flags and the posters you've got.
    
    Right.
    
    We to deal with this the best way possible for everyone is to take your details and have you on your way. All right. If you don't provide us with your name and address, those are necessities for us to arrest you for offences against Terrorism Act for being part of the proscribed group. So if we can take your name and your address and then you could be on your way, we can leave it as that and if you go about something a different day, it'll be dealt with a different day. But for today now on this busy roundabout rush hour where we don't want an accident to happen, that's how we want to deal with it is take your details and leave. Failing that, you may render yourself liable to arrest in order in order for us to deal with it. Does that make sense?
    
    Does that make sense? That makes sense.
    
    That's how we want to deal with this. Okay. Because we want to keep traffic going, not have to leave you in a shell. If we can take your details, someone will be in touch and they'll deal with it another way. Are you happy to give us your details and then be on your way so that we can deal with it that way?
    
    On the basis of that, then I will give you those [cut]
    
    ‘Cos you slowed the traffic down, distracting drivers, right? Do this in the city center.
    
    If I do this in a city center, are you guys going to come back and tell me that I'm now potentially committing a terrorist offence? Because I have the flag of another country again. Have a flag of another country.
    
    You have a green flag. Can it just be a flag? You don't just have a flag. No, I don't. I have Free Gaza, and a statement of fact.
    
    We also have to think about your safety.
    
    In the middle of the day, plenty of people have been very supportive. plenty of people don't go free for your cause.
    
    I know plenty of people do.
    
    I'm sure there is. But you’re going to come across a person who will take exception to it.
    
    [cut]
    
    So, I'll tell you the offences that are we've been asked to deal with. So, you've got to belong or profess to belong to a proscribed organization.
    
    Which I have not.
    
    It's a personal [inaudible] and you haven't. All right.
    
    And you can't prove that I do.
    
    Invite support for a proscribed organization. You haven't. Express an opinion or belief that is supportive of a proscribed organization. Now, that you have done.
    
    Well, that's ridiculous. You're going to end up arresting half the country.
    
    I didn't make the legislation. That is the legislation.
    
    That’s extremely broad.
    
    [inaudible] your details. And that's why if we take your details and we tell you to leave, we'll tell you not to carry on this down the high street. The high street is the better place because it's not a road. We're going to tell you not to, take your details and be on our way. If you're found down the high street later on, you'll probably find yourself arrestable to that fence because you've been warned not to.
    
    So, I can never do this again in Canterbury, in Kent.
    
    It’s an offence.
\end{quote}

Some very dim people will construct conspiracy theories, perhaps antisemitic, on this basis (probably couched in terms of ‘Zionists’ and so on). But the common pattern in each is that the police unthinkingly go after easy targets. If they want to find people who support Palestine Action more reliably, they will surely find more in the larger \textsc{psc} marches—but these tend to be so well-attended that it’s not (to them) worth the bother (cf when Gideon Falter was stopped from walking through a crowd because of his kippah). Similarly, when Niyak Ghorbani calls Hamas terrorists at \textsc{psc} marches, the police \href{https://www.standard.co.uk/news/london/metropolitan-police-hamas-is-terrorist-placard-b1144341.html}{arrest} him, because he’s the easier target. Only when there is substantial political pressure will they act otherwise (e.g. during the race riots) and often act with considerable courage, but, of course, such political pressure can only opaquely be exercised by the Home Secretary, in the name of ‘operational independence’ (which only really exists when convenient).

In a better world, there would be a substantial segment of public opinion—including the leaders of the \textsc{psc}—against all these arrests and threatened arrests. But we get the plods we deserve, and if we can’t build a coalition from Gideon Falter to the \textsc{psc} demanding a sensible approach to public order policing, it’ll be the best efforts of such articulate officers as the ones quoted above that’ll suffice.