Ariel Rubinstein…

Israel, Ariel Rubinstein
===
…on ‘the catastrophe, after which everything will, probably, be different’ \parencite{rubinstein2015a}.
\begin{quote}
    The flagship issues for the relgious-Zionist camp are unity of the people and unity of the land. They are getting all of the land via the settlements and the occupation. And they’re getting all of the people for free because most of us are addicted to the mantras and refrain from speaking the truth: almost a half-century of occupation and settlement is separating us, the “faithful of the Jewish people”, from them, the “faithful of the Land of Israel.”

    Our lot is not with them. There is no place for false groveling or shows of reconciliation. This false unity is enabling the settlers to achieve all their goals. Instead of searching for common ground, we should be declaring our glaring differences. The Israeli right also only understands credible threats. The religious-Zionist public will be compelled to reexamine its positions if it is made to understand that spiritual redemption is not around the corner, and that fealty of the land is going to shatter the unity of the people.

    And if this last force is of no help, all that will be left to do (for me, at least) is to stay here, watch with horror the oppression of the Palestinians, watch as a society whose values I don’t respect takes shape and await—with trepidation—the catastrophe, after which everything will, probably, be different.
\end{quote}

…on ‘correct Zionist action’ \parencite{rubinstein2010}.
\begin{quote}
    These young people are here simply because this is their home and in any other place they would feel that they do not belong. They are here not because of the tax benefits but because of the familiar brand of instant coffee. They are here because they are actually dying to do reserve duty, even if only to have the opportunity to refuse to serve in the territories.
\end{quote}

…on an ‘internal commission of inquiry’ \parencite{rubinstein2006}.
\begin{quote}
    Here in Israel, this is the season for commissions of inquiry. And I feel the need to appoint an internal commission of inquiry. On this committee, there will be one member (me), one witness (me) and one person under scrutiny (me). The committee will focus on the question: How did I ever become an enthusiastic supporter of Amir Peretz, the Labour Party leader, during Israel’s election campaign earlier this year?
\end{quote}